---
title: "Homework 6"
author: "Marisa Sobel"
date: "11/16/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
library(mgcv)
library(broom)
library(boot)
library(modelr)

set.seed(1)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

# Problem 1

### DATA IMPORT AND TIDY

```{r, collapse=TRUE, message=FALSE}
wp_homicide_data = 
  read_csv("./data/data-homicides-master/homicide-data.csv", na = c("", "Unknown", "NA")) %>% 
  mutate(city_state = str_c(city, state, sep = ", ")) %>% 
  filter(!city_state %in% c("Dallas, TX", "Pheonix, AZ", "Kansas City, MO", "Tulsa, AL")) %>% 
  filter(victim_race != "NA") %>% 
  filter(victim_age != "NA") %>% 
  filter(victim_sex != "NA") %>% 
  mutate(
    victim_race = as_factor(case_when(
      victim_race == "White" ~ "white", 
      victim_race != "white" ~ "non-white")), 
    victim_race = fct_relevel(victim_race, "white"),
    victim_sex = as_factor(tolower(victim_sex)), 
    victim_sex = fct_relevel(victim_sex, "female"), 
    victim_age = as.numeric(victim_age),
    outcome = as_factor(case_when(
      disposition == "Closed by arrest" ~ "resolved",
      disposition == "Closed without arrest" ~ "unresolved", 
      disposition == "Open/No arrest" ~ "unresolved"))) %>% 
  select(-uid, -reported_date, -victim_first, -victim_last, -lat, -lon, -city, -state, -disposition)

wp_homicide_data
```

The original dataset had 52,169 entries. `r 52169-49411` entries were removed from cities that did not have race data in full ("Dallas, TX", "Pheonix, AZ", "Kansas City, MO", "Tulsa, AL"). `r 49411 - 47969` entries were removed for missing race information, `r 47969 - 47578` entries were removed for missing age information, and `r 47578 - 47525` entries were removed for missing sex information.  

Victim race is coded as binary as white and non-white, with white as the reference category.  
Victim sex is coded as binary as female and male, with female as the refrence category.  
The homicide outcome is coded as binary as resolved and unresolved, where resolved includes "Closed by arrest" and unresolved includes "closed without arrest" and "open/no arrest". 

### BALTIMORE, MD  

Run `glm` (resolved/unresolved) for Baltimore, MD comparing non-white victims to white victims and accounting for sex and age. 

```{r}
# baltimore glm model 
glm_baltimore = 
  wp_homicide_data %>% 
  filter(city_state == "Baltimore, MD") %>% 
  glm(outcome ~ victim_age + victim_sex + victim_race, data = ., family = binomial)

# get OR and CIs
tidy(glm_baltimore, conf.int = TRUE) %>% 
  mutate(OR = exp(estimate), 
         CI_low = exp(conf.low), 
         CI_high = exp(conf.high)) %>% 
  select(term, OR, CI_low, CI_high) %>% 
  filter(term == "victim_racenon-white") %>% 
  knitr::kable(digits = 3)
```

For the city of Baltimore, MD, the odds of solving homicides among non-whites is 0.441 (95% CI: 0.312, 0.62) times the odds of solving homicides among whites. Conversly, the odds of solving homicides among whites is `r round(1/0.441, digits = 2)` times the odds of solving homicides among non-whites in Baltimore, MD. 

### ALL CITIES  
#### GLM (resolved/unresolved)

Run `glm` (resolved/unresolved) for all cities comparing non-white victims to white victims and accounting for sex and age. 

```{r}
# all cities glm model 
glm_all_cities = 
  wp_homicide_data %>% 
  group_by(city_state) %>% 
  nest() %>% 
  mutate(
    model = map(data, ~tidy(glm(outcome ~ victim_age + victim_sex + victim_race, 
                                data = ., family = binomial), conf.int = TRUE))) %>% 
  select(-data) %>% 
  unnest()
  
# get OR and CIs
or_all_cities = 
  glm_all_cities %>% 
  mutate(OR = exp(estimate), 
         CI_low = exp(conf.low), 
         CI_high = exp(conf.high)) %>% 
  select(city_state, term, OR, CI_low, CI_high) %>% 
  filter(term == "victim_racenon-white") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  arrange(city_state, desc(OR))

or_all_cities
```

Use the `or_all_cities` dataframe to plot ORs and CIs for all cities. 

### ALL CITIES
#### Plot - ORs + CIs 

```{r}
or_all_cities %>% 
  ggplot(aes(x = city_state, y = OR)) +
  geom_hline(aes(yintercept = 1), color = "red") +
  geom_point() +
  geom_errorbar(aes(ymin = CI_low, ymax = CI_high)) + 
  coord_flip() + 
  labs(
    x = "OR (95% CI)", 
    y = "City, State", 
    title = "Solved homicides, comparing non-white victims to white victims")
```

After adjusting for victim age and sex, the plot above illustrates the ORs (95% CI) of a homicide being solved comparing non-white victims to white victims. Most cities have higher odds of homicides being solved if the victim is white (OR < 1), except for Tampa, Birmigham, and Durham. Durham has the largest CI, and there seems to be a trend of larger CIs the closer the OR gets to 1, the null value for ORs. 

# Problem 2

### DATA IMPORT AND TIDY

```{r, message=FALSE, collapse=TRUE}
# load data and transform factors
birthweight_data = 
  read_csv("./data/birthweight.csv") %>% 
  mutate(
    babysex = factor(babysex), 
    frace = factor(frace), 
    malform = factor(malform), 
    mrace = factor(mrace))

birthweight_data

# check for missing data
birthweight_data %>% 
  is.na() %>% summary()
```

There are `r nrow(birthweight_data)` entries in the birthweight dataset. All of the original subjects have all data; no entry was dropped. 

### PROPOSED REGRESSION MODEL - BIRTHWEIGHT

I propose a linear model as birthweight is a continuous outcome. Based on previous knowledge, infant birthweight is affected by gestational age, mother's weight gain during pregnancy, and cigarette smoking. I am also including mother's BMI before pregnancy to account for differences in body type. I am expecting increases in gestational age and mother's weight gain to have a positive effect on birthweight, whereas increased smoking should have a decreased effect. I am refraining from adding in other demographic parameters at the moment as I do not want to overcomplicate the model. 

```{r, collapse=TRUE}
# naming model
bwt_lm = lm(bwt ~ gaweeks + ppbmi + wtgain + smoken, data = birthweight_data)
  
# model output
bwt_lm_output = 
  birthweight_data %>% 
  lm(bwt ~ gaweeks + ppbmi + wtgain + smoken, data = .) %>% 
  tidy(conf.int = TRUE) %>% 
  select(term, estimate, conf.low, conf.high)

birthweight_data %>% 
  lm(bwt ~ gaweeks + ppbmi + wtgain + smoken, data = .) %>% 
  glance()

bwt_lm_output %>% 
  knitr::kable(digits = 3)
```

As expected, increased gestational age, mother's weight gain, and mother's BMI before pregnancy all resulted in positive beta coefficients, whereas increased smoking had a negative effect on birthweight. 

### RESIDUALS PLOT

```{r, message=FALSE}
# add residuals and predicitions to `birthweight_data`
residual_plot = 
  birthweight_data %>% 
  modelr::add_residuals(bwt_lm) %>% 
  modelr::add_predictions(bwt_lm) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  geom_smooth() +
  labs(
    x = "Predictions", 
    y = "Residuals", 
    title = "Proposed model residuals vs. fitted values"
  )

residual_plot
```

The residuals decrease at first, increase in the thick of the plot, and decrease again. However, there is no observable trend. 

### MODEL COMPARISON

1. Length at birth and gestational age as predictors (main effects only). 
2. Head circumference, length, sex, and all interactions (including the three-way interaction) between these. 

```{r, collapse=TRUE}
# model 1 comparison 
model_compare_1 = lm(bwt ~ blength + gaweeks, data = birthweight_data)

model_compare_1 %>% tidy()

model_compare_1 %>%  glance()

# model 2 comparison 
model_compare_2 = lm(bwt ~ bhead + blength + babysex + bhead*blength 
                     + bhead*babysex + blength*babysex + bhead*blength*babysex, 
                     data = birthweight_data)

model_compare_2 %>% tidy()

model_compare_2 %>%  glance()
```

### CROSS VALIDATION

```{r}
#cross_validation = 
  
```








